\Rpackage{MSnbase} development

### 2011-03-21 inst/doc/MSnbase-demo.Rnw
MSnSet now extends eSet and has MIAPE expressionData slot; MIAPE extended with MIAPE-MS slots

<<environment,cache=FALSE,echo=FALSE>>=
library(MSnbase)
@

@@

\Rpackage{MSnbase} aims are providing a reproducible research framework to the proteomics data analysis. 
It should allow researcher to easily mine mass spectrometry data, explore the data and its statistical properties 
and visually display these.

@@

Raw data is stored in an \Robject{MSnExp} object, that stores all the spectra in an experiment, 
as defined by one or multiple raw data files.

\Rpackage{MSnbase} imports raw MS data stored in \texttt{mzXML} 
format \citep{Pedrioli2004}. More formats, notably \texttt{mzML} \citep{Martens2010}, 
will be added at a later stage. 
Either one of MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.

<<MSnExp,cache=FALSE,echo=TRUE>>=
library(MSnbase)
filename <- dir(system.file(package="MSnbase",dir="extdata"),
                full.name=TRUE,pattern="mzXML$")
print(filename)
experiment <- readMzXMLData(filename, msLevel = 2, verbose=FALSE)
experiment
@

@@

As illustrated above, showing the experiment textually displays it's content: 
\begin{itemize}
  \item Information about the raw data, i.e. the spectra.
  \item Specific information about the experiment processing\footnote{%%
    this part will be automatically updated when the object is 
    modified with it's methods, as illustrated later} and package version.
    This slot can be accessed with the \Rfunction{processingData} method.
  \item Other meta data, including experimental phenotype, 
    file name(s) used to import the data, protocol data, 
    information about features (individual spectra here), 
    experiment data and proteomics data. Most of these are implemented as 
    in the \Robject{eSet} class and are described in more details in their 
    respective manual pages. See also \texttt{?pSet} and \texttt{MSnExp}
    and references therein for additional background information.

    The proteomics-specific meta data in slot \Robject{proteomicsData} is a 
    new addition from the \Rpackage{MSnbase} package.
    It contains MIAPE data (Minimum Information About a Proteomics Experiment, \citep{Taylor2007}) 
    and can be accessed with the \Rfunction{proteomicsData} method. 
    See \texttt{?MIAPE} for more details.
\end{itemize}

@@

Attributes of individual spectra or of all spectra of an experiment can be 
accessed with their respective methods: \Rfunction{precursorCharge} for the 
precursor charge, \Rfunction{rtime} for the retention time, \Rfunction{mz} 
for the MZ values, \Rfunction{intensity} for the inensities, \ldots see the 
respecitve class manuals for more details.

@@

Reporter ions are defined with the \Robject{ReporterIons} class. 
Specific peaks of interested are defined by a MZ value, a with 
around the expected MZ and a name (and optionally a colour for plotting, 
see section \ref{sec:plotting}). \Robject{ReporterIons} instances 
are required to quantitate reporter peaks in \Robject{MSnExp} 
experiments. Instances for the most commonly used isobaric tags like 
iTRAQ 4-plex and 8-plex and TMT tags are already defined in 
\Rpackage{MSnbase}. See \texttt{?ReporterIons} for details about how 
to generate new \Robject{ReporterIons} objects.

@@

Spectra can be plotted individually or as part of (subsetted) experiments 
with the \Robject{plot} method. Full spectra can be plotted 
(using \texttt{full=TRUE}), specific reporter ions of interest 
(by specifing with reporters with \texttt{reporters=iTRAQ4} for instance) 
or both (figure \ref{fig:spectrum-plot}).

\begin{figure}[!ht]
<<spectrum-plot,fig=T,echo=T,tikz=T,external=T,width=5,height=5>>=
plot(sp,reporters=iTRAQ4,full=TRUE)
@ 
\caption{Raw MS2 spectrum with detail about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

It is also possible to plot all spectra of an experiment (figure \ref{fig:msnexp-plot}). 
Lets start by extracting all spectra that have the sample precursor MZ value than \Robject{sp} 
into a separate experiment, using the \Rfunction{extractPrecSpectra} method. 
It takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.

<<extractPrecSpec,echo=T,cache=F>>=
exp2 <- extractPrecSpectra(experiment,precursorMz(experiment)["X43"])
exp2
@

We see that \Sexpr{length(exp2)} spectra have the same MZ value. 
There two spectra can be represented together by plotting the 
\Robject{MSnExp} object (figure \ref{fig:msnexp-plot}).

@@

There are several methods implemented to perform basic data manipulation. 
Low intensity peaks can be set to 0 with the \Rfunction{removePeaks} method 
from spectra or whole experiments. The intensity below which peaks are removed 
is defined by the \Robject{t} parameter. \texttt{t} can be specified directly 
as a numeric. The default value is the character \texttt{"min"}, that will remove 
all peaks equal to the lowest intensity peaks in the spectrum.

@@

Unlike the name might suggest, the \Rfunction{removePeaks} method does 
not remove peaks from the spectrum. To effectively remove 0 intensity 
peaks from spectra, and reduce the size of the data set, one can use 
the \Rfunction{clean} method.

@@

Another useful manipulation method is \Rfunction{trimMz}, that taken as 
parameters and \Robject{MSnExp} (or a \Robject{Spectrum}) and a numeric 
\Robject{mzlim}. MZ values smaller to \texttt{min(mzlim)} 
or greater then \texttt{max(mzmax)} are discarded. This method is 
particularely useful when one wants to concentrate on a specific MZ range, 
as for reporter ions quantification, as is generally results in substantial 
reduction of data size. Compare the size of the full trimmed experiment to 
the original \Sexpr{sz} Mb.

@@

@@

The \Rfunction{quantify} method returns \Robject{MSnSet} objects, that 
extend the well-known \Robject{ExpressionSet} objects defined in the 
\Rpackage{Biobase} package. The assay data is a matrix of dimensions 
$n \times m$ matrix, where $m$ is the  number of features/spectra 
originally in the \Robject{MSnExp} used as parameter in 
\Robject{quantify} and $m$ is the number of reporter ions. 
The meta data is directly inherited from the \Robject{MSnExp} instance.

@@

A \Robject{MSnSet} object is mean to be compatible with further downstream 
packages for data normalisation and statistical analysis. 
Below is an example of applying 
variance stabilization normalisation \citep{Huber2002} 
to the iTRAQ reporter intensities \citep{Karp2010}.

<<vsn,echo=TRUE,cache=FALSE>>=
library(vsn)
qnt.vsn <- vsn2(qnt)
@

@@

<<cvdata,echo=FALSE,cache=FALSE>>=
sd1 <- apply(log2(exprs(qnt))+10,1,sd)
mn1 <- apply(log2(exprs(qnt))+10,1,mean)
cv1 <- sd1/mn1
sd2 <- apply(exprs(qnt.vsn)+10,1,sd)
mn2 <- apply(exprs(qnt.vsn)+10,1,mean)
cv2 <- sd2/mn2
library(zoo)
## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
## rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
##
## Calling directly rollapply.zoo to make it zoo_1.6-4 compatible.
## The above requires zoo >= 1.7-0, which is as of 15 March 2011
## not yet available on CRAN (only on r-forge).
rmed1 <- zoo:::rollapply.zoo(cv1,7,function(x) median(x,na.rm=TRUE))
rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
@

\begin{figure}[!ht]
<<cv-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=5>>=
plot(order(mn1),cv1,col="steelblue",pch=19,
     ylab="CV",xlab="Rank mean")
points(order(mn2),cv2,col="red",pch=19)
grid()
lines(seq(1,70,by=(70/length(rmed1))),rmed1,col="steelblue",lwd=2)
lines(seq(1,70,by=(70/length(rmed2))),rmed2,col="red"      ,lwd=2)
## legend("topright",
##        legend=c("log2","vsn","running\nmedian"),
##        pch=c(19,19,NA),
##        col=c("steelblue","red","black"),
##        lwd=c(0,0,2),
##        bty="n",
##        cex=.7)
@ 
\caption{CV versus signal intensity comparison for log2 (red) and vsn (red) transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

>>>>

<<environment,cache=FALSE,echo=FALSE>>=
library("MSnbase")
@

@@

\Rpackage{MSnbase} aims are providing a reproducible research framework to proteomics data analysis. 
It should allow researcher to easily mine mass spectrometry data, explore the data and its statistical properties 
and visually display these.

@@

Raw data is contained in \Robject{MSnExp} objects, that stores all the spectra of an experiment, 
as defined by one or multiple raw data files.

\Rpackage{MSnbase} imports raw MS data stored in \texttt{mzXML} 
format \citep{Pedrioli2004} with the function \Rfunction{readMzXMLData} function. 
More formats, notably \texttt{mzML} \citep{Martens2010}, will be added at a later stage. 
Either MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.

<<MSnExp,cache=FALSE,echo=TRUE>>=
library("MSnbase")
filename <- dir(system.file(package="MSnbase",dir="extdata"),
                full.name=TRUE,pattern="mzXML$")
print(filename)
experiment <- readMzXMLData(filename, msLevel = 2, verbose=FALSE)
experiment
@

@@

As illustrated above, showing the experiment textually displays it's content: 
\begin{itemize}
  \item Information about the raw data, i.e. the spectra.
  \item Specific information about the experiment processing\footnote{%%
    this part will be automatically updated when the object is 
    modified with it's \textit{ad hoc} methods, as illustrated later} and package version.
    This slot can be accessed with the \Rfunction{processingData} method.
  \item Other meta data, including experimental phenotype, 
    file name(s) used to import the data, protocol data, 
    information about features (individual spectra here) and
    experiment data. Most of these are implemented as 
    in the \Robject{eSet} class and are described in more details in their 
    respective manual pages. See \texttt{?MSnExp}
    and references therein for additional background information.

    The experiment meta data associated with an \Robject{MSnExp} experiment
    is of class \Robject{MIAPE}. It stores general information about the 
    experiment as well as MIAPE (Minimum Information About a Proteomics Experiment, 
    \citep{Taylor2007, Taylor2008}) information.
    This meta-data can be accessed with the \Rfunction{experimentData} method. 
    When available, a summary of MIAPE-MS data can be printed with the \Rfunction{msInfo} method.
    See \texttt{?MIAPE} for more details.
\end{itemize}

@@

Attributes of individual spectra or of all spectra of an experiment can be 
accessed with their respective methods: \Rfunction{precursorCharge} for the 
precursor charge, \Rfunction{rtime} for the retention time, \Rfunction{mz} 
for the MZ values, \Rfunction{intensity} for the inensities, \ldots see the 
\Robject{Spectrum}, \Robject{Spectrum1} and \Robject{Spectrum2} manuals for more details.

@@

Reporter ions are defined with the \Robject{ReporterIons} class. 
Specific peaks of interest are defined by a MZ value, a with 
around the expected MZ and a name (and optionally a colour for plotting, 
see section \ref{sec:plotting}). \Robject{ReporterIons} instances 
are required to quantify reporter peaks in \Robject{MSnExp} 
experiments. Instances for the most commonly used isobaric tags like 
iTRAQ 4-plex and 8-plex and TMT tags are already defined in 
\Rpackage{MSnbase}. See \texttt{?ReporterIons} for details about how 
to generate new \Robject{ReporterIons} objects.

@@

Spectra can be plotted individually or as part of (subsetted) experiments 
with the \Robject{plot} method. Full spectra can be plotted 
(using \texttt{full=TRUE}), specific reporter ions of interest 
(by specifing with reporters with \texttt{reporters=iTRAQ4} for instance) 
or both (see figure \ref{fig:spectrum-plot}).

\begin{figure}[!ht]
<<spectrum-plot,fig=T,echo=T,tikz=T,external=T,width=5,height=5>>=
plot(sp,reporters=iTRAQ4,full=TRUE)
@ 
\caption{Raw MS2 spectrum with details about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

It is also possible to plot all spectra of an experiment (figure \ref{fig:msnexp-plot}). 
Lets start by extracting all spectra that have the same precursor MZ value than \Robject{sp} 
into a separate experiment, using the \Rfunction{extractPrecSpectra} method. 
This method takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.

<<extractPrecSpec,echo=T,cache=F>>=
exp2 <- extractPrecSpectra(experiment,precursorMz(sp))
exp2
@

We see that \Sexpr{length(exp2)} spectra have the same MZ value. 
These can be visulaised together by plotting the 
\Robject{MSnExp} object, as illustrated on figure \ref{fig:msnexp-plot}.

@@

There are several methods implemented to perform basic data manipulation. 
Low intensity peaks can be set to 0 with the \Rfunction{removePeaks} method 
from spectra or whole experiments. The intensity threshold below which peaks are removed 
is defined by the \Robject{t} parameter. \Robject{t} can be specified directly 
as a numeric. The default value is the character \texttt{"min"}, that will remove 
all peaks equal to the lowest non null intensity in any spectrum. 
We observe the effect of the \Rfunction{removePeaks} method by comparing 
total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{tic} method
before (object \Robject{sp}) and after (object \Robject{experiment}). 
Figure \ref{fig:spectrum-plot2} illustrated spectrum \texttt{X43} after low intensity 
peaks removal.

@@

Unlike the name might suggest, the \Rfunction{removePeaks} method does 
not actually remove peaks from the spectrum; they are set to 0. This 
can be checked using the \Rfunction{peaksCount} method, that returns 
the number of peaks (including 0 intensity peaks) in a spectrum.
To effectively remove 0 intensity 
peaks from spectra, and reduce the size of the data set, one can use 
the \Rfunction{clean} method.

@@

Another useful manipulation method is \Rfunction{trimMz}, that takes as 
parameters and \Robject{MSnExp} (or a \Robject{Spectrum}) and a numeric 
\Robject{mzlim}. MZ values smaller then \texttt{min(mzlim)} 
or greater then \texttt{max(mzmax)} are discarded. This method is 
particularely useful when one wants to concentrate on a specific MZ range, 
as for reporter ions quantification, and generally results in substantial 
reduction of data size. Compare the size of the full trimmed experiment to 
the original \Sexpr{sz} Mb.
\label{timMz-example}

@@

As can be seen above, all processing performed on the experiment is recorded 
and displayed as integral part of the experiment object.

@@

The \Rfunction{quantify} method returns \Robject{MSnSet} objects, that 
extend the well-known \Robject{eSet} class defined in the 
\Rpackage{Biobase} package. \Robject{MSnSet} instances are very similar 
to \Robject{ExpressionSet} objects, except for the experiment meta-data 
that captures MIAPE specific information.
The assay data is a matrix of dimensions 
$n \times m$ matrix, where $m$ is the  number of features/spectra 
originally in the \Robject{MSnExp} used as parameter in 
\Robject{quantify} and $m$ is the number of reporter ions, that 
can be accessed with the \Rfunction{exprs} method.
The meta data is directly inherited from the \Robject{MSnExp} instance.

@@

A \Robject{MSnSet} object is meant to be compatible with further downstream 
packages for data normalisation and statistical analysis. 
Below is an example of applying 
variance stabilization normalisation \citep{Huber2002} 
to the iTRAQ reporter intensities \citep{Karp2010}.

<<vsn,echo=TRUE,cache=FALSE>>=
library("vsn")
qnt.vsn <- vsn2(exprs(qnt))
@

@@

<<cvdata,echo=FALSE,cache=FALSE>>=
sd1 <- apply(log2(exprs(qnt))+10,1,sd)
mn1 <- apply(log2(exprs(qnt))+10,1,mean)
cv1 <- sd1/mn1
sd2 <- apply(exprs(qnt.vsn)+10,1,sd)
mn2 <- apply(exprs(qnt.vsn)+10,1,mean)
cv2 <- sd2/mn2
library("zoo")
## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
## rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
##
## Calling directly rollapply.zoo to make it zoo_1.6-4 compatible.
## The above requires zoo >= 1.7-0, which is as of 15 March 2011
## not yet available on CRAN (only on r-forge).
rmed1 <- zoo:::rollapply.zoo(cv1,7,function(x) median(x,na.rm=TRUE))
rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
@

\begin{figure}[!ht]
<<cv-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=5>>=
plot(order(mn1),cv1,col="steelblue",pch=19,
     ylab="CV",xlab="Rank mean")
points(order(mn2),cv2,col="red",pch=19)
grid()
lines(seq(1,70,by=(70/length(rmed1))),rmed1,col="steelblue",lwd=2)
lines(seq(1,70,by=(70/length(rmed2))),rmed2,col="red"      ,lwd=2)
## legend("topright",
##        legend=c("log2","vsn","running\nmedian"),
##        pch=c(19,19,NA),
##        col=c("steelblue","red","black"),
##        lwd=c(0,0,2),
##        bty="n",
##        cex=.7)
@ 
\caption{CV versus signal intensity comparison for log2 (blue) and vsn (red) transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

### 2011-03-21 inst/doc/MSnbase-development.Rnw
MSnSet now extends eSet and has MIAPE expressionData slot; MIAPE extended with MIAPE-MS slots

This document is not replacement for the individual manual pages, that  
slots of the \Rpackage{MSnbase} classes. It is a centralised high-level 
description of the package design.

@@

This virtual class aims coordinate contain mass spectrometry 
data, i.e spectra, and meta data. It is based on the \Robject{eSet} for genomic 
data. The main difference with \Robject{eSet} that the \Robject{assayData} slot is an 
environment containing any number of \Robject{Spectrum} instances (see section \ref{sec:spectrum}).

Two new slots are introduced, namely \Robject{proteomicsData}, that contains 
\Robject{MIAPE} instances (see section \ref{sec:miape}) and 
\Robject{processingData}, that contains \Robject{MSnProcess} instances 
(see section \ref{sec:msnprocess}). The \Robject{annotation} slot has not 
been implemented, as no prior feature annotation is known in shotgun proteomics.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{MSnExp}: a class MS experiments}
\label{sec:msnexp}

\Robject{MSnExp} extends \Robject{pSet} to 
an is used to store MS experiments. It does not add any new slots 
to \Robject{pSet}. Accessors and setters are all inherited from 
\Robject{pSet} and new ones should be implemented for \Robject{pSet}. 
Methods that manipulate actual data in experiments are implemented 
for \Robject{MSnExp}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{MSnSet}: a class quantitative proteomics data}

This class stores quantitation data and meta data 
after running \Robject{quantify} on an \Robject{MSnExp} object. The 
quantitative data is in form of a $n \times m$ matrix, where $m$ is the 
number of features/spectra originally in the \Robject{MSnExp} used as 
parameter in \Robject{quantify} and $m$ is the number of reporter ions 
(see section \ref{sec:reporterions}). This thus prompted to directly extend 
the \Robject{ExpressionSet} and add two proteomics-specific annotation slots
introduced in the \Robject{pSet} class, namely \Robject{proteomicsData} for 
\Robject{MIAPE} objects (see section \ref{sec:miape}) and 
\Robject{processingData} for objects of class \Robject{MSnProcess} 
(see section \ref{sec:msnprocess}). The \Robject{annotation} slot, 
inherited from \Robject{eSet} is not used.

@@

This class aims at recording specific manipulations 
applied to \Robject{MSnExp} or \Robject{MSnSet} instances. The \Robject{processing} 
slot is a \Robject{character} vector that describes major processing. Most 
other slots are of class \Robject{logical} that indicate whether the data has been 
centroided, smoothed, \ldots although many of the functionality is not implemented yet.
Any new processing that is implemented should be documented logged here.

It also documents the raw data file from which the data originates (\Robject{files} 
slot) and the \Rpackage{MSnbase} version that was installed when the \Robject{MSnProcess} 
instance was originally created.

@@

The Minimum Information About a Proteomics 
Experiment \citep{Taylor2007} \Robject{MIAPE} class is currently
only composed by a free text \Robject{description} attribute. This 
class will be expanded in future versions. 
Note that the MIAPE standard information that overlaps with MIAME 
requirements should directly be documented in the \Robject{experimentData} 
slots (see also note on page \pageref{misc:whymiame}).

Raw data is currently imported from \texttt{mzXML} files \citep{Pedrioli2004} 
using the \Rfunction{xcms:::rampRawData} 
and \Rfunction{xcms:::rampRawDataMSn} functions from the 
\Rpackage{xcms} package \citep{Smith2006}.
These functions do not give access to the meta data. New importer functions are 
under development (see for instance \Rpackage{mzR}\footnote{%%
  \url{https://github.com/sneumann/mzR/blob/master/DESCRIPTION}}) that will 
hopefully give programmatic access to meta data stored in the data file to 
populate the \Robject{MIAPE} object.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{Spectrum} \it{et al.}: classes for MS spectra}
\label{sec:spectrum}

@@

The iTRAQ and TMT (or any other peak of interest) 
are implemented with \Robject{ReporterIons}, 
that essentially defines an expected MZ position for the peak and a width around 
this value as well a names for the reporters.

@@

\paragraph{Why this MIAME slot?} \label{misc:whymiame} While it would have been 
possible to transfer all data stored in \Rpackage{Biobase}'s \Robject{MIAME} 
to a new \Robject{MIAPE} class and use the latter for \Robject{experimentData} slots in 
the \Robject{pSet} class, it would not have been possible to directly transfer this 
to \Robject{MSnSet} instances, as \Robject{MSnSet} classes directly inherit 
from the \Robject{ExpressionSet}, whose \Robject{experimentData} slot must be of class 
\Robject{MIAME}.

>>>>

This document is not a replacement for the individual manual pages, that  
document the slots of the \Rpackage{MSnbase} classes. It is a centralised high-level 
description of the package design.

@@

This virtual class is the main container for mass spectrometry 
data, i.e spectra, and meta data. It is based on the \Robject{eSet} implementation for genomic 
data. The main difference with \Robject{eSet} is that the \Robject{assayData} slot is an 
environment containing any number of \Robject{Spectrum} instances (see section \ref{sec:spectrum}).

One new slot is introduced, namely \Robject{processingData}, 
that contains one \Robject{MSnProcess} instance  (see section \ref{sec:msnprocess}). 
 and the \Robject{experimentData} slot is now expected to contain
\Robject{MIAPE} data (see section \ref{sec:miape}).
The \Robject{annotation} slot has not 
been implemented, as no prior feature annotation is known in shotgun proteomics.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{MSnExp}: a class for MS experiments}
\label{sec:msnexp}

\Robject{MSnExp} extends \Robject{pSet} to store MS experiments. 
It does not add any new slots 
to \Robject{pSet}. Accessors and setters are all inherited from 
\Robject{pSet} and new ones should be implemented for \Robject{pSet}. 
Methods that manipulate actual data in experiments are implemented 
for \Robject{MSnExp} objects.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{MSnSet}: a class for quantitative proteomics data}

This class stores quantitation data and meta data 
after running \Robject{quantify} on an \Robject{MSnExp} object. The 
quantitative data is in form of a $n \times m$ matrix, where $m$ is the 
number of features/spectra originally in the \Robject{MSnExp} used as 
parameter in \Robject{quantify} and $m$ is the number of reporter ions 
(see section \ref{sec:reporterions}).

This prompted to keep a similar implementation as 
the \Robject{ExpressionSet} class, while adding the proteomics-specific annotation slot
introduced in the \Robject{pSet} class, namely 
\Robject{processingData} for objects of class \Robject{MSnProcess} 
(see section \ref{sec:msnprocess}).

The \Robject{MSnSet} class extends the virtual \Robject{eSet} class 
to provide compatibility for \Robject{ExpressionSet}-like behaviour. 
The experiment meta-data in  \Robject{experimentData} is 
also of class \Robject{MIAPE} (see section \ref{sec:miape}).
The \Robject{annotation} slot, inherited from \Robject{eSet} is not used.

@@

This class aims at recording specific manipulations 
applied to \Robject{MSnExp} or \Robject{MSnSet} instances. The \Robject{processing} 
slot is a \Robject{character} vector that describes major processing. Most 
other slots are of class \Robject{logical} that indicate whether the data has been 
centroided, smoothed, \ldots although many of the functionality is not implemented yet.
Any new processing that is implemented should be documented and logged here.

It also documents the raw data file from which the data originates (\Robject{files} 
slot) and the \Rpackage{MSnbase} version that was in use when the \Robject{MSnProcess} 
instance, and hence the \Robject{MSnExp}/\Robject{MSnSet} objects, were originally created.

@@

The Minimum Information About a Proteomics 
Experiment \citep{Taylor2007, Taylor2008} \Robject{MIAPE} class 
describes the experiment, including contact details, 
information about the mass spectrometer and control 
and analysis software.

Raw data is currently imported from \texttt{mzXML} files \citep{Pedrioli2004} <
using the \Rfunction{xcms:::rampRawData} 
and \Rfunction{xcms:::rampRawDataMSn} functions from the 
\Rpackage{xcms} package \citep{Smith2006}.
These functions do not give access to the meta data. New importer functions are 
under development (see for instance \Rpackage{mzR}\footnote{%%
  \url{https://github.com/sneumann/mzR/blob/master/DESCRIPTION}}) that will 
hopefully give programmatic access to meta data stored in the data file to 
populate the \Robject{MIAPE} object.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{Spectrum} \textit{et al.}: classes for MS spectra}
\label{sec:spectrum}

@@

The iTRAQ and TMT (or any other peak of interest) 
are implemented \Robject{ReporterIons} instances, 
that essentially defines an expected MZ position for the peak and a width around 
this value as well a names for the reporters.

@@

%% \paragraph{Why this MIAME slot?} \label{misc:whymiame} While it would have been 
%% possible to transfer all data stored in \Rpackage{Biobase}'s \Robject{MIAME} 
%% to a new \Robject{MIAPE} class and use the latter for \Robject{experimentData} slots in 
%% the \Robject{pSet} class, it would not have been possible to directly transfer this 
%% to \Robject{MSnSet} instances, as \Robject{MSnSet} classes directly inherit 
%% from the \Robject{ExpressionSet}, whose \Robject{experimentData} slot must be of class 
%% \Robject{MIAME}.
